---
title: "Introduction to rGEO"
author: "Zhuoer Dong"
date: "`r Sys.Date()`"
output: prettydoc::html_pretty
vignette: >
  %\VignetteIndexEntry{Introduction to rGEO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

options(tibble.print_min = 6, tibble.max_extra_cols = 1)
```

```{r}
library(rGEO)
```

# quick start

You can find raw data at https://www.ncbi.nlm.nih.gov/geo/browse/ ^[search for 
    `GSE***`, download files from the `Download family` section at the bottom.],
    or use the following handy tools:
    
```{r}
gse_soft_ftp('GSE51280')

gse_matrix_ftp('GSE51280')
```
    
Then you can read them:

```{r}
# expression matrix file
read_gse_matrix(system.file('extdata/GSE51280_series_matrix.txt.gz', package = 'rGEO'))

# SOFT (chip annotation) file
read_gse_soft(system.file('extdata/GSE51280_family.soft.gz', package = 'rGEO'))
```

# A little deeper

Let's go a little deeper to see how this works. 

The package mainly contains functions for   
   1. telling the ftp path to GSE raw data: see `gds_ftp()` and `gpl_soft_ftp()`  
   1. reading these files, including expression matrix file and SOFT (chip 
        annotation) file, see below.  

## expression matrix file

reading expression matrix file is quiet straightforward, you just need to locate
    where the information lies, and `readr::read_tsv()` would take care of the 
    rest for you.

```{r eval=FALSE}
read_gse_matrix
```

```{r results='asis', echo=FALSE}
cat('```r\n')
read_gse_matrix
cat('```\n')
```


## SOFT (chip annotation) file

SOFT (chip annotation) file is a lot more harder to deal with. In short, you 
    have to dig useful information first:
    
```{r}
parse_gse_soft(system.file('extdata/GSE51280_family.soft.gz', package = 'rGEO'), F)
```

and convert that to a uniform format. Almost the whole package does the latter, 
    refer to `vignette('probe2symbol')` for more details. 

By the way, for both `parse_gse_soft()` and `read_gse_soft()`, you can add 
    `verbose = T` to make the process more clear.



